---
title: "Unit 2 Paper"
author: "Rachit Agarwal"
date: "2/28/2019"
output:
  word_document: default
  html_document: default
---
### Overview: 
This file gathers information on the msa data. We summarize the missing values from the data. We do exploratory analysis on the data. Finally, we implement and test linear regession models using many differnt methods for assesment. 

### Apendex 1 Detail of Statistical Models:
  Conceptual proofs
  a.  Find log(Y/N) = B0 + B1log(N) from Y = c(N^b) where c>0 and b>1. 
      Y = cN^b
      Y/ N = cN^(b-1) 
      log(Y/N) = log(cN^(b-1))
      log(Y/N) = log(c) + (b-1)log(N)
      since c>0 and b>1 we have B0 and B1>0
  b.	Find log(Y) = B0 + (1+B1)log(N) from log(Y/N) = B0 + B1log(N):
      log(Y/N) = B0 + B1log(N)
      log(Y) – log(N) = B0 + B1log(N) 
      log(Y) – log(N) – B1log(N) = B0
      log(Y) – (1+B1)log(N) = B0 
      log(Y) = B0 + (1+B1)log(N)
For our first hypothesis, when taking into account shares of the economy deriving from finance and ict and how they relate to GMP it produces a better linear model than the relationship between GMP and population size. 

Additionally, when taking into account shares of the economy deriving from finance, ict and prof.tech and how theyrelate to GMP it produces a better linear model than the relationship between GMP and population size. 

### External Requirments
```{r}
library(tidyverse)
library(ggplot2)
library(knitr)
library(dplyr)
```
### Apendex 2 Exploratory Analyis:
```{r cars}
# read in missing data
msadata=read.csv("http://dept.stat.lsa.umich.edu/~bbh/s485/data/gmp-2006.csv")
## read in holdout sample
holdout=read.csv("http://dept.stat.lsa.umich.edu/~bbh/s485/data/gmp-2006-holdout.csv")
## add all neccary variables to the data
GMP = as.numeric(msadata$pop)*as.numeric(msadata$pcgmp)
msadata=data.frame(msadata,GMP)

#number of missing values 
sum(is.na(msadata$finance))
sum(is.na(msadata$prof.tech))
sum(is.na(msadata$ict))
sum(is.na(msadata$management))
sum(is.na(msadata))
## filter complete cases for the four sectors
finance <- mean(complete.cases(dplyr::select(msadata, finance)))
prof.tech <- mean(complete.cases(dplyr::select(msadata, prof.tech)))
ict <- mean(complete.cases(dplyr::select(msadata, ict)))
management <- mean(complete.cases(dplyr::select(msadata, management)))
#filter complete cases by pairs
mean(complete.cases(dplyr::select(msadata, finance, prof.tech)))
mean(complete.cases(dplyr::select(msadata, finance, ict)))
mean(complete.cases(dplyr::select(msadata, finance, management)))
mean(complete.cases(dplyr::select(msadata, prof.tech, ict)))
mean(complete.cases(dplyr::select(msadata, prof.tech, management)))
mean(complete.cases(dplyr::select(msadata, ict, management)))
#complete cases for entire dataset
mean(complete.cases(dplyr::select(msadata,finance,prof.tech,ict,management)))
##final dataset#

##scaterplots
ggplot(data=msadata,mapping = aes(x=pop,y=GMP)) + geom_point(alpha=0.5)+geom_smooth(se=FALSE)
ggplot(data=msadata,mapping = aes(x=pop,y=log(GMP))) + geom_point(alpha=0.5)+geom_smooth(se=FALSE)
ggplot(data=msadata,mapping = aes(x=log(pop),y=GMP)) + geom_point(alpha=0.5)+geom_smooth(se=FALSE)
ggplot(data=msadata,mapping = aes(x=log(pop),y=log(GMP))) + geom_point(alpha=0.5)+geom_smooth(se=FALSE)

##scaterplots with colors corresponding to other variables
ggplot(data=msadata,mapping = aes(x=log(pop),y=log(GMP),color=finance)) + geom_point(alpha=0.5)+geom_smooth(se=FALSE)
ggplot(data=msadata,mapping = aes(x=log(pop),y=log(GMP),color=prof.tech)) + geom_point(alpha=0.5)+geom_smooth(se=FALSE)
ggplot(data=msadata,mapping = aes(x=log(pop),y=log(GMP),color=ict)) + geom_point(alpha=0.5)+geom_smooth(se=FALSE)
ggplot(data=msadata,mapping = aes(x=log(pop),y=log(GMP),color=management)) + geom_point(alpha=0.5)+geom_smooth(se=FALSE)

```
### Apendex 3 Fitting the power law model:
```{r}
## fitting model
power_law<-lm(log(msadata$pcgmp)~log(msadata$pop))
summary(power_law)
## verifying with residuals
plot(log(msadata$pop),resid(power_law))
## in sample loss
in_sample_loss	= mean(resid(power_law)^2)
in_sample_loss
## cross validation
cv.lm <- function(data, formulae, nfolds = 5) 
{
  data <- na.omit(data)
  formulae <- sapply(formulae, as.formula)
  n <- nrow(data)
  fold.labels <- sample(rep(1:nfolds, length.out = n))
  mses <- matrix(NA, nrow = nfolds, ncol = length(formulae))
  colnames <- as.character(formulae)
  for (fold in 1:nfolds) {
    test.rows <- which(fold.labels == fold)
    train <- data[-test.rows, ]
    test <- data[test.rows, ]
    for (form in 1:length(formulae)) {
      current.model <- lm(formula = formulae[[form]], data = train)
      predictions <- predict(current.model, newdata = test)
      test.responses <- eval(formulae[[form]][[2]], envir = test)
      test.errors <- test.responses - predictions
      mses[fold, form] <-mean(test.errors^2)
    }
  }
  return(colMeans(mses))
}
cross_valid<-"log(msadata$pcgmp)~log(msadata$pop)"
cv = cv.lm(msadata,cross_valid,5)
cv
```
From the proofs derived above, we know that log(GMP) = log(c) + log(b-1)log(pop), therefore form the model summary of the power law we can determine what the values of c and b are. 
The coeficent for the intercept is 8.796 so 8.796 log(c) and c = 6607.76. 
The coeficent for log(pop) is 0.123 which is equal to b-1 so b = 1.123.
This is compatible with the supra-linear power law because c>0 and b>1.

Accoridng to the residual plot, we see that there is no indication of a pattern and therefore there is a porper fit for the power law and we can use the summary function and beleive the standard eror it provides which is 0.238 on 242 degrees of freedom.  

### Apendex 4 Fitting the assement of alternate models:
```{r}
## now we have to delete missing values for finance,ict and prof.tech from our dataset ##
analyis_data = drop_na(msadata,finance,ict,prof.tech)
## alternate models
original_mod<-lm(log(analyis_data$pcgmp)~log(analyis_data$pop))
summary(original_mod)
fin_pop<-lm(log(analyis_data$pcgmp)~analyis_data$finance)
summary(fin_pop)
tech_pop<-lm(log(analyis_data$pcgmp)~analyis_data$ict)
summary(tech_pop)
mod4<-lm(log(analyis_data$pcgmp)~analyis_data$ict+analyis_data$finance)
summary(mod4)
mod5<-lm(log(analyis_data$pcgmp)~analyis_data$ict+analyis_data$finance+analyis_data$prof.tech)
summary(mod5)

##in-sample loss 
original_in_sample_loss	= mean(resid(original_mod)^2)
fin_in_sample_loss	= mean(resid(fin_pop)^2)
tech_in_sample_loss	= mean(resid(tech_pop)^2)
mod4_in_sample_loss	= mean(resid(mod4)^2)
mod5_in_sample_loss	= mean(resid(mod5)^2)

original_in_sample_loss
fin_in_sample_loss 
fin_in_sample_loss
tech_in_sample_loss
mod4_in_sample_loss
mod5_in_sample_loss

## cross validataon ##
cross_valid1<-"log(analyis_data$pcgmp)~log(analyis_data$pop)"
cross_valid2<-"log(analyis_data$pcgmp)~analyis_data$finance"
cross_valid3<-"log(analyis_data$pcgmp)~analyis_data$ict"
cross_valid4<-"log(analyis_data$pcgmp)~analyis_data$ict+analyis_data$finance"
cross_valid5<-"log(analyis_data$pcgmp)~analyis_data$ict+analyis_data$finance+analyis_data$prof.tech"
cv1 = cv.lm(analyis_data,cross_valid1,5)
cv1
cv2 = cv.lm(analyis_data,cross_valid2,5)
cv2
cv3 = cv.lm(analyis_data,cross_valid3,5)
cv3
cv4 = cv.lm(analyis_data,cross_valid4,5)
cv4
cv5 = cv.lm(analyis_data,cross_valid5,5)
cv5
```
### Apendex A Nested alternative model:
```{r}
## nested model for best altetrnative model
nested_model<- lm(log(analyis_data$pcgmp)~analyis_data$ict+analyis_data$finance+analyis_data$prof.tech+log(analyis_data$pop))
summary(nested_model)
## anova test comparing nestred model with alternative model 
anova(nested_model,mod5)
```

### Apendex B Holdout sample: 
```{r}
# B #
## drop missing values from hold out dataset
hold_data = drop_na(msadata,finance,ict,prof.tech)
# add yi or orginal value to holdout dataset
hold_data = data.frame(hold_data,log(hold_data$pcgmp))


#predictions on holdout data 
predictions_original<-predict(original_mod, newdata = hold_data)
predictions_fin<-predict(fin_pop, newdata = hold_data)
predictions_tech<-predict(tech_pop, newdata = hold_data)
predictions_mod4<-predict(mod4, newdata = hold_data)
predictions_mod5<-predict(mod5, newdata = hold_data)

## calculate test errors and MSE for each model 
holdout_errors_original = hold_data$log.hold_data.pcgmp - predictions_original
MSE_original<-mean((holdout_errors_original)^2)
MSE_original

holdout_errors_fin = hold_data$log.hold_data.pcgmp - predictions_fin
MSE_fin<-mean((holdout_errors_fin)^2)
MSE_fin

holdout_errors_tech = hold_data$log.hold_data.pcgmp - predictions_tech
MSE_tech<-mean((holdout_errors_tech)^2)
MSE_tech

holdout_errors_mod4 = hold_data$log.hold_data.pcgmp - predictions_mod4
MSE_mod4<-mean((holdout_errors_mod4)^2)
MSE_mod4

holdout_errors_mod5 = hold_data$log.hold_data.pcgmp - predictions_mod5
MSE_mod5<-mean((holdout_errors_mod5)^2)
MSE_mod5

```

